---
title: "COVID Forecasting"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Data and Visualize


```{r}
covid_data = read.csv(file = "./covid-data-us-national.csv")

head(covid_data)
```

Visualizing forecasting target

```{r}
library(ggplot2)
# Basic line plot with points
ggplot(data=covid_data, aes(x=epiweek, y=death_jhu_incidence, group=1)) +
  geom_line()+
  geom_point()
```

Visualizing features/predictors, e.g. cases

```{r}
library(ggplot2)
# Basic line plot with points
ggplot(data=covid_data, aes(x=epiweek, y=covid_cases, group=1)) +
  geom_line()+
  geom_point()
```


## Data pre-processing

Data pre-processing: scaling input

```{r}
min_max_norm <- function(x) {
    (x - min(x)) / (max(x) - min(x))
  }
covid_data = as.data.frame(lapply(covid_data[4:7], min_max_norm))
time.series = covid_data$death_jhu_incidence
head(covid_data)
```

## Modeling with ARIMA 

We fit ARIMA and predict for epiweeks 202030 to 202043

```{r}
pred_func <- function(week, order, k_ahead=1){
    train.series = time.series[1:week-k_ahead]
    model = arima(train.series, order = order)
    pred = predict(model, k_ahead)
    return(pred$pred)
}

arima.preds = c()
order = c(2,2,2)  # can be tuned
for (w in 30:40){
    arima.preds = append(arima.preds, pred_func(w, order))
}

arima.preds
```
## Modeling with Off-the-shelf Neural Network

Now we will use neural networks to predict future forecasts from features of current week.

```{r}
library(neuralnet)
library(dplyr)

make.dataset <- function(ahead=1) {
    full.dataset = as.data.frame(covid_data[(10-ahead):(40-ahead),])
    labels = covid_data$death_jhu_incidence[10:40]
    full.dataset$labels = labels
    return(full.dataset)
}

dataset = make.dataset()
head(dataset)
```

Predict for epiweeks 202030 to 202043

```{r}
nn_train <- function(w, hidden = c(10, 40)) {
    model = neuralnet(labels~death_jhu_incidence+mobility+totalTests+covid_cases,
                      data=dataset[1:(w-10),], hidden=hidden, linear.output=T)
    return(model)
}

nn.preds = c()
for(w in 30:40){
    m = nn_train(w)
    nn.preds = append(nn.preds, predict(m, dataset[w-10+1,]))
}

nn.preds
```

## Evaluation

```{r}
ground.truths = covid_data$death_jhu_incidence[30:40]
ground.truths

arima.rmse = sqrt(mean((arima.preds-ground.truths)^2))
nn.rmse = sqrt(mean((nn.preds-ground.truths)^2))
print(c("Arima RMSE:", arima.rmse))
print(c("NN RMSE", nn.rmse))
```

Our ARIMA is better than an off-the-shelf simple neural network. Can we do better?